\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex

### 2023-05-25 paper-seb-supplementary-material.tex
fixing final bugs

%\rfoot{\supplement\ \thepage/\SupplementalLastPage}
\lfoot{\today}
\cfoot{}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\thispagestyle{empty}

@@

%\makeatletter
%% saves name-value pairs in both the supplementary material .aux file and the info for the main file
%\def\remembernamevalue#1#2{%
%	\immediate\write\@auxout{\@backslashchar gdef\@backslashchar #1{#2}}%
%	\immediate\write\infofileForMain{\@backslashchar gdef\@backslashchar #1{#2}}%
%}
%\makeatother

@@

% writing to the aux file, this file generates labels like
% \newlabel{supplementary-best-practice}{{2}{5}{Software engineering best practice}{section.2}{}}
% whereas the main file using IEEEtrans class generates
% \newlabel{table-summary}{{I}{5}} or \newlabel{jvs-policy}{{V}{4}} etc

@@

% start supplementary material's page counters to continue on from the paper's page numbering
\setcounterpageref{page}{LastPage}
\addtocounter{page}{1} % otherwise we start with the same page that the main paper ended on (we want to be on the next page:-)!

%\setcounter{section}{0}
\setcounter{section}{\continueSectionNumbers}
\def\n#1{{\sf\bfseries ({#1})}}

\newdimen \www 
\setbox0=\hbox{\n{viii}~}
\www=\wd0 \advance \www by .4em
\def\nnInContents#1#2{\hbox to \www {\hfill\n{#1}~\hskip .4em} #2\\}

\setbox0=\hbox{\sfdefault\large(00)}
\newdimen\howwide \howwide=\wd0 \howwide=2.5em

\def\sec#1#2#3{\setbox0=\hbox{\n{#1}~~}\leftskip \wd0
%\begin{flushleft}
\subsubsection*{\setbox0=\hbox{\n{#1}~~}\hskip -\wd0\copy0 \bfseries
\immediate\write\minicontentsfile{\bslash nnInContents{#1}{#3}}%
{#2}\starredsubsectioncontents{\hbox to \howwide{\hfill (#1)} #3}}
%\end{flushleft}
\leavevmode
\vskip 1ex
\par\leftskip 0in}

\def\highlightsec#1#2#3{\sec{#1}{{#2}}{{#3}}}
\def\highlightmysubsection#1{\subsection{{#1}}\subsectioncontents{{#1}}}
\def\changemysubsection#1{\subsection{{#1}}\subsectioncontents{{#1}}}
\def\changemysection#1{\vskip 2ex\section{{#1}}\sectioncontents{{#1}}}
\def\highlightmysection#1{\section{{#1}}\sectioncontents{{#1}}}
\def\mysection#1{\section{#1}\sectioncontents{#1}}
\def\mysubsection#1{\subsection{#1}\subsectioncontents{#1}}

@@

\def\sectioncontents#1{\immediate\write\contentsfile{\bslash contents{\arabic{section} #1}{\thepage}}}
\def\subsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{\alph{subsection}\ #1}{\thepage}}}
\def\starredsubsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{#1}{\thepage}}}

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent\small
\input generated/supplementary.toc
}

\newwrite\contentsfile
\immediate\openout\contentsfile=generated/supplementary.toc

\newpage
\begin{center}
\emph{Pagination, section, and citation numbering continues from the main paper}
\end{center}
\mysection{Further issues for Software Engineering Boards (SEBs)}

\highlightmysubsection{Brief definition}
Software Engineering Boards, henceforth SEBs, will be used to help and assure that critical code, including epidemic modeling, is of high standard, to provide assurance for scientific papers, Government public health and other policies, etc, that the code used is of appropriate quality for its intended uses.

@@

\mysubsection{Relationships of SEBs to Ethics Boards}

@@

\changemysubsection{SEBs are necessary but not sufficient}
The main paper provides evidence and argues that SEBs (or equivalent) are necessary to help improve the quality of science, specifically science relying, explicitly or implicitly, on tools or methods based in software.

@@

\mysection{Software engineering best practice}
\label{supplementary-best-practice}
\mysubsection{Introduction and standard references}

@@

\mysubsection{Essential components of best practice}
Software Engineering includes the following topics, which are discussed at more length below:
\vskip 2ex

\vbox{\noindent
      \input generated/minicontents.tex
      }

\newwrite\minicontentsfile
\immediate\openout\minicontentsfile=generated/minicontents.tex

\sec{0}{Without defining requirements, not enough skilled effort will be put into designing and implementing reliable software --- or excess effort will be wasted}{Requirements}
It is not always necessary to program well if the code to be produced is for fun, experimenting, or for demonstrations. On the other hand, if code is intended for life-critical applications, then it is worth putting more engineering effort into it. The first step of Software Engineering, then, is to assess the requirements, specifically the reliability requirements of the code that is going to be produced.

@@

\highlightsec{1}{Without formal methods, there is no rigorous and checked specification of a program, so nobody --- including its developers --- will know exactly what it is supposed to do}{Formal methods}

@@

\sec{2}{Without defensive programming, any errors --- in data, code, hardware, or in use --- will go unnoticed and be uncorrected}{Defensive programming} Defensive programming is based on a range of methods, including error checking, independent calculation (using multiple implementations written by independent programmers), assertions, regression testing, etc. Notoriously, what are often unconsciously dismissed as trivial concerns frequently lead to the hardest to diagnose errors, such as buggy handling of ``well-known, trivial'' things like numbers \cite{numerals}. The great advantage of defensive programming is that it detects, and may be able to recover from, bugs that have been missed earlier in the development process (such as typos in the code). Defensive programming requires professional training to be used effectively, for example it is not widely known that some choices of programming language make defensive programming unnecessarily hard \cite{heedless}.

@@

\highlightsec{3}{Using inappropriate programming languages undermines reliability}{Using dependable programming languages}
Many popular languages are popular because they are easy to use, which is not the same as being reliable to use. The fewer constraints a language imposes, the easier it \emph{seems\/} to be to program in, but the lack of constraints means the language cannot provide the checks stricter languages do. C, for instance, which is one of the languages widely used for modeling \cite{tweet,plos}, is not a good choice for a reliable programming language --- it has many intrinsic weaknesses that are well-known to professionals, but which frequently trap inexperienced programmers. (This is not the place for a review of bad programming languages, for which see \cite{heedless}, but Excel is even worse than C\@.)

@@

\sec{4}{Version control and open source organizes and helps software development}{Open source and version control}
It is appreciated that the models may change and be adapted as new data and insights become available. Changing models makes it even harder to ensure that they are correct, and thus emphasizes the relevance of the core message this paper: we have to find ways to make computer models more reliable, inspectable, and verifiable. Version control keeps a record of what code was used when, and enables reconstruction of earlier versions of code that has been used. Version control is supported by many tools (such as Git, Subversion, etc).

@@

\sec{5}{Without professional testing, there is no acceptable evidence that a program works under real conditions}{Rigorous testing}
In poorly-run software development it is very easy to miss bugs, because the flawed thinking that inserted bugs in the code is going to be the same flawed thinking with the same misconceptions that tries to detect them. Rigorous testing includes methods like fault injection. Here, the idea is that if testing finds no bugs, that may be because the testing is not rigorous enough rather than that the program actually has no bugs. Fault injection inserts random bugs, and then testing gives statistical insights into the number of bugs in a program (depending on how many deliberate bugs it successfully finds).

@@

\highlightsec{6}{Without documentation and record keeping, nobody --- least of all the programmer --- knows what code is supposed to do or how to get it to do it}{Good documentation and record keeping}
Documentation covers internal documentation (how code works), developer (how to include it in other programs), configuration (how to configure and compile the code in different environments), external documentation (how the code is used), and help (documentation available while using the program).

@@

\sec{7}{If code is not usable, even if it is ``correct'' it will not be used and interpreted correctly}{Usability}
\globalLabel{supplementary-material-makefiles}
Usability is an important consideration: \cite{hci1,hci2} is the program usable by its intended users so they can obtain correct results? Often the programmers developing code know it so well they misjudge how easy it will be for anyone else to use it --- this is a very serious problem for the lone programmer (possibly working in another country) supporting a research team. Usability is especially important when programs are to be used by other researchers and by non-programmers, including epidemiologists.

@@

\sec{8}{Without using existing solutions (libraries, APIs, etc) reinventing code merely reinvents bugs}{Reusing quality solutions}
Reusing quality code (mathematical functions, database operations, user interface features, connectivity, etc) avoids having to develop it oneself, saves time and avoids the risks of introducing new bugs. The more code that is reused, the more likely many people will have contributed to improving it --- for example, reusing a standard database package will provide Atomicity, Consistency, Isolation, and Durability (so-called ACID properties) without any further work (nor even needing to understand what useful guarantees these basic properties ensure).

@@

\sec{9}{Poor programmers often fix bugs rather than the causes of bugs: complexity and obfuscation}{Simplicity} 
When a program doesn't quite do what is wanted, it is tempting to add more features or variables, or to treat the problem as an ``exception'' and program around it --- which inserts more code and, almost certainly, more bugs. This way lies over-fitting, a problem familiar from statistics (and machine learning). Programs can be made over-complex and they can then do anything; an over-complex program may seem correct by accident. Instead, the hallmarks of good science are that of parsimony and simplicity; if a simple program can do what is needed it is more likely to be correct. A simpler program is easier to prove correct, easier to program, and easier to debug. A special case of needing simplicity is when fixing bugs: instead of fixing bugs one at a time, one should be fixing the \emph{reasons\/} why the bugs have happened. Generally, when bugs are fixed, programmers should determine \emph{why\/} the bugs occurred, and thence repair the program more strategically.

\highlightsec{10}{International standards have been developed to support critical software development}{Compliance with standards}

@@

\sec{11}{Effective multidisciplinary teamwork is essential because no individual has the capacity to develop non-trivial reliable software}{Effective multidisciplinary teamwork}

@@

\highlightsec{12}{Computing technologies are advancing rapidly, and best practice in Software Engineering is continually evolving}{Continuous Professional Development (CPD)}

@@

\highlightsec{13}{Other factors \ldots}{Security and other factors}
Of course, there are many other factors to be considered for the professional development of critical code, such as using appropriate methods to ensure cybersecurity \cite{security-engineering,cyber-cacm}, particularly while also being able to up- and download secure updates.

For pandemic modeling specifically, understanding the limitations of numerical methods (in particular, how numerical methods are affected by the choice of programming language and style of programming) is critical.\footnote{{For example, one of the surveyed paper's code \cite{example-numerical-error} uses literal numbers at far too high a precision for the chosen language to be able to represent correctly (conformant implementations use IEEE 754 double precision 64-bit floating point). Such an error typically has an undefined impact on results, and unfortunately is easy to overlook as the program almost certainly ignores the error when running. The error belies misunderstandings in programming which may have wider effects, such as consequences of relying on the precision being higher than it is.}} Hamming \cite{hamming} is considered a classic, but there is a huge choice available.

@@

\highlightsec{14}{Human Factors are the foundation of everything}{Software is a human activity}

@@

Following the Dunning-Kruger Effect \cite{dunning-kruger,dunning-kruger-numeracy}, programmers over-estimate their programming skills because they do not have the skills to recognize their lack of knowledge, specifically in the present case, knowledge of basic Software Engineering. Dunning and Kruger go on to say,

@@

\mysection{Code, data, and publication}
\globalLabel{on-code-data-publication}

@@

\mysubsection{When magic numbers become magic code}
Data often controls the flow of code. For example, data summarizing patients may include their gender, but the program processes males and females differently. Then data becomes code.

@@

\mysubsection{When data is code}
{Many computer programs blur the simplistic code/data distinctions deliberately, to create virtual machines. Data is then run on the virtual machine as program. Many programs provide standard features to do this, such as LISP's and JavaScript's \texttt{eval} functions. Henderson's book \cite{henderson} builds an elegant Pascal program to run \emph{any\/} LISP program as data, and then shows that the LISP program can run itself running other programs, so it is now its own code and \emph{its\/} data --- despite being purely data to the Pascal program. There are numerous advantages to doing this, including: the Pascal program is not just reading data, but structured data that must conform to the rules of LISP; the LISP running itself runs faster than the original Pascal running LISP, even though the Pascal virtual machine is still doing it in the recursive case; LISP is a much more powerful language than Pascal, so a virtual machine can be used to escape the barriers of a limited implementation such as Pascal. In short, any distinctions between code and data are impossible to maintain.}

@@

\mysubsection{Exploiting code as data for more reliable science}
In the present paper, we knowingly built on this blur between data and code, a special case of \RAPstarp. However, what we did was not unusual except in our explicit and rigorous approach to managing and summarizing data reliably in the paper.

@@

\mysubsection{When data is text: Exploiting code for reliable publication}

@@

\mysubsection{Data and polynomials used in the paper}
The data and polynomials used in the main paper's section \ref{over-fit}, and illustrated in the paper's figure \ref{fig-overfit}, is presented below, as generated in \LaTeX\ by the same \emph{Mathematica\/} notebook that generated section \ref{over-fit}. Of course, in the paper itself, the data was not necessary to make the point, but if there is any need to replicate it or otherwise scrutinize the argument in the paper --- as there would be for more complex arguments in typical papers --- the data used and results are shown here. In a similar way, data and computed results could be presented and made available in other scientific papers.

@@

\highlightmysubsection{Comparing conventional and RAP approaches}
\globalLabel{RAP-diagrams}
The similarities and differences between the conventional copy-and-paste approach to filling in data and diagrams in publications, the improved systematic RAP process, and using notebooks (such as \emph{Mathematica\/} or Jupyter), are illustrated in the sequence of schematics of figures \ref{fig:basic}, \ref{fig:excel}, \ref{fig:rap}, and \ref{fig:notebook}.

@@

\changemysection{The Spiegelhalter trustworthiness questions}
\globalLabel{supplementary-Spiegelhalter-section}

@@

\def\questionsection#1{\changemysubsection{#1}}
\def\question#1{\item \emph{#1\/}}
\def\sequestion#1{\begin{itemize}\raggedright\item[$\blacktriangleright$] \textbf{#1}\end{itemize}}

\questionsection{How trustworthy are the numbers?}
\newcounter{resumeCounter}

@@

\sequestion{If the claims are exploratory, weaker standards of coding can be used; if the claims are a basis for critical decisions, then there should be evidence of using appropriate Software Engineering (such as defensive programming) to provide appropriate confidence in the results claimed.}
\setcounter{resumeCounter}{\value{enumi}}
\end{enumerate}
\questionsection{How trustworthy is the source?}
\begin{enumerate}
\setcounter{enumi}{\value{resumeCounter}}
\question{How reliable is the source of the story?} Consider the possibility of a biased source with conflicts of interest, and check publication is independently peer-reviewed. Ask yourself, `Why does this source want me to hear this story?'

@@

\setcounter{resumeCounter}{\value{enumi}}
\end{enumerate}
\questionsection{How trustworthy is the interpretation?}
\begin{enumerate}
\setcounter{enumi}{\value{resumeCounter}}
\question{How does the claim fit with what else is known?} Consider the context, appropriate comparators, including historical data, and what other studies have shown, ideally in a meta-analysis.

@@

\highlightmysection{A pilot survey of computational science}
\globalLabel{survey-section}

@@

The \RAPstar\ approach cannot enforce the truth of such as summary, but it (and the opportunity to read and review it easily) very significantly increases the chances that the summary is correct and up to date. For example, if refactoring leads to an option being deleted, then it will also disappear \emph{with no further work\/} from the table above. Also, since the summary was proof-read at the same time as proof-reading this \supplement, following \RAPstar\ also increases the chances that any errors or functionality omissions or issues in the makefile have been detected and corrected. data.js

\highlightmysubsection{Selected journal case studies}

@@

In 2020 the \emph{Journal of Vascular Surgery\/} published a research paper \cite{jvs1}, which had to be retracted on ethical grounds \cite{jvs2,jvs3}: it was a na\"\i ve study and the editorial process was unaware of digital norms. Notably, the paper fails to provide access to its anonymized data (with or without qualification), and fails to define the data anonymization algorithm, and also fails to even mention the code that it developed and used to perform its study. The journal's data policy is itself very weak (the authors ``should consider'' including a footnote to offer limited access to the data) and, despite basic statistics policies, it has no policy at all for code (see  section \ref{supplementary-journal-policies-section}). Ironically, the retracted article \cite{jvs1} is still online (as of August 2020) with no reference to any editorial statement to the effect that it has been retracted, despite this being trivial --- and necessary --- to achieve in the widely-accessed online medium.

@@

\highlightmysubsection{Pilot paper sample}
A sample of \plural{\dataN}{recent paper} covering a broad range of science were sampled from the leading journals \journalBreakdown.

@@

%\newcount \total
%\total=\dataN
%\multiply \total by 4 % three reviewers, one editor
%\advance \total by \countAuthors
%\the\total\
%leading scientists assuming authors, 1 editor, and 3 reviewers per paper, but not counting acknowledged colleagues et al, covering  (plus appendices and supporting material) in total.
%
\caption{Overview of the peer-reviewed paper sample.}
\label{supplement-table-overview}
\end{table*}

@@

\caption{Summary of survey results.}
\label{supplement-table-summary}
\end{table*}

@@

\highlightmysubsection{Summary of results}
\label{summary-of-results}
The sample selection criteria necessarily identified scientific research with Software Engineering contributions.

@@

\highlightmysubsection{Current code policies of sampled journals}
\globalLabel{supplementary-journal-policies-section}

It is noteworthy that none of the journals sampled permit any reliable style of managing data in published papers, such as described above in sections \ref{detecting-against-error} and \ref{restrictive-policies}. In particular, for all the papers that had accessible code, the code included explicit (and relevant) data that was not archived \emph{as\/} data in the journal repositories.

Note that journal policies were first accessed on 29 July 2020, close after the period covered by the pilot survey, thus presumably fairly closely reflecting the policies in use for the papers in the survey on the dates when they were each submitted. Unfortunately for the survey, the journals do not make clear what exact policies were applied to each paper when the papers were submitted (on the other hand, the survey shows that policies are not rigorously enforced).
\\

\newdimen\tabwidth
\tabwidth=\textwidth
\advance \tabwidth by -4em
\def\specifyurl{\vskip 3mm
\begingroup\raggedleft\scriptsize
} 
\def\specifyurletc#1{\endgroup\\
\hbox{ }\hfill\begin{tabular}{|p{\tabwidth}@{}}
#1
\end{tabular}\vskip .3cm
}

{\sf
\noindent\textbf{Extract from \emph{Royal Society Open Science\/} author guidelines} \\
--- It is a condition of publication that authors make the primary data, materials (such as statistical tools, protocols, software) and code publicly available. These must be provided at the point of submission for our Editors and reviewers for peer-review, and then made publicly available at acceptance. [\ldots] As a minimum, sufficient information and data are required to allow others to replicate all study findings reported in the article. Data and code should be deposited in a form that will allow maximum reuse. As part of our open data policy, we ask that data and code are hosted in a public, recognized repository, with an open licence (CC0 or CC-BY) clearly visible on the landing page of your dataset.}

\specifyurl
\url{royalsociety.org/journals/authors/author-guidelines/#data}
\specifyurletc{Since first accessed 29 July 2020, the policy has been revised (undated, accessed 2 February 2022) but retains the same principles; full policy now available via a DOI \cite{RS-policy}. The policy still retains an emphasis on data accessibility, and continues a lack of awareness that code and data are equivalent and often mixed (see section \ref{on-code-data-publication}).}

\noindent\textbf{Extract from \emph{Nature Digital Medicine\/} author guidelines}\\
{\sf --- A condition of publication in a Nature Research journal is that authors are required to make materials, data, code, and associated protocols promptly available to readers without undue qualifications. [\ldots] A condition of publication in a Nature Research journal is that authors are required to make unique materials promptly available to others without undue qualifications.}

>>>>

\lfoot{\today}
\cfoot{}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\thispagestyle{empty}

@@

@@

@@

% start supplementary material's various counters to continue on from the paper's page numbering
\setcounter{section}{\continueSectionNumbers}
\setcounter{table}{\continueTableNumbers}
\setcounter{figure}{\continueFigureNumbers}
\setcounter{footnote}{\continueFootnoteNumbers}
\setcounter{page}{\continuePageNumbers}
\addtocounter{page}{1} % otherwise we start with the same page that the main paper ended on (we want to be on the next page:-)!

@@

\newpage
\begin{center}
\emph{Section, figure, table, citation, and pagination numbering continue from the main paper}
\end{center}

\setcounter{tocdepth}{3}

\tableofcontents
\newpage

\section{Further issues for Software Engineering Boards (SEBs)}

Software Engineering Boards, henceforth SEBs, will be used to help and assure that critical code, including epidemic modeling, is of high standard, to provide assurance for scientific papers, Government public health and other policies, etc, that the code used is of appropriate quality for its intended uses.

@@

\subsection{Relationships of SEBs to Ethics Boards}

@@

\subsection{SEBs are necessary but not sufficient}
The main paper provides evidence and argues that SEBs (or equivalent) are necessary to help improve the quality of science, specifically science relying, explicitly or implicitly, on tools or methods based in software.

@@

\section{Software engineering best practice}
\label{supplementary-best-practice}
%\subsection{Introduction and standard references}

@@

%\subsection{Essential components of best practice}
Software Engineering includes the following topics:

\subsection{Requirements}
It is not always necessary to program well if the code to be produced is for fun, experimenting, or for demonstrations. On the other hand, if code is intended for life-critical applications, then it is worth putting more engineering effort into it. The first step of Software Engineering, then, is to assess the requirements, specifically the reliability requirements of the code that is going to be produced.

@@

\subsection{Formal methods}

@@

\subsection{Defensive programming} 
Defensive programming is based on a range of methods, including error checking, independent calculation (using multiple implementations written by independent programmers), assertions, regression testing, etc. Notoriously, what are often unconsciously dismissed as trivial concerns frequently lead to the hardest to diagnose errors, such as buggy handling of ``well-known, trivial'' things like numbers \cite{numerals}. The great advantage of defensive programming is that it detects, and may be able to recover from, bugs that have been missed earlier in the development process (such as typos in the code). Defensive programming requires professional training to be used effectively, for example it is not widely known that some choices of programming language make defensive programming unnecessarily hard \cite{heedless}.

@@

\subsection{Using inappropriate programming languages undermines reliability}
Many popular languages are popular because they are easy to use, which is not the same as being reliable to use. The fewer constraints a language imposes, the easier it \emph{seems\/} to be to program in, but the lack of constraints means the language cannot provide the checks stricter languages do. C, for instance, which is one of the languages widely used for modeling \cite{tweet,plos}, is not a good choice for a reliable programming language --- it has many intrinsic weaknesses that are well-known to professionals, but which frequently trap inexperienced programmers. (This is not the place for a review of bad programming languages, for which see \cite{heedless}, but Excel is even worse than C\@.)

@@

\subsection{Open source and version control}
It is appreciated that the models may change and be adapted as new data and insights become available. Changing models makes it even harder to ensure that they are correct, and thus emphasizes the relevance of the core message this paper: we have to find ways to make computer models more reliable, inspectable, and verifiable. Version control keeps a record of what code was used when, and enables reconstruction of earlier versions of code that has been used. Version control is supported by many tools (such as Git, Subversion, etc).

@@

\subsection{Rigorous testing}
In poorly-run software development it is very easy to miss bugs, because the flawed thinking that inserted bugs in the code is going to be the same flawed thinking with the same misconceptions that tries to detect them. Rigorous testing includes methods like fault injection. Here, the idea is that if testing finds no bugs, that may be because the testing is not rigorous enough rather than that the program actually has no bugs. Fault injection inserts random bugs, and then testing gives statistical insights into the number of bugs in a program (depending on how many deliberate bugs it successfully finds).

@@

\subsection{Good documentation and record keeping}
Documentation covers internal documentation (how code works), developer (how to include it in other programs), configuration (how to configure and compile the code in different environments), external documentation (how the code is used), and help (documentation available while using the program).

@@

\subsection{Usability}
\globalLabel{supplementary-material-makefiles}
Usability is an important consideration: \cite{hci1,hci2} is the program usable by its intended users so they can obtain correct results? Often the programmers developing code know it so well they misjudge how easy it will be for anyone else to use it --- this is a very serious problem for the lone programmer (possibly working in another country) supporting a research team. Usability is especially important when programs are to be used by other researchers and by non-programmers, including epidemiologists.

@@

\subsection{Reusing quality solutions}
Reusing quality code (mathematical functions, database operations, user interface features, connectivity, etc) avoids having to develop it oneself, saves time and avoids the risks of introducing new bugs. The more code that is reused, the more likely many people will have contributed to improving it --- for example, reusing a standard database package will provide Atomicity, Consistency, Isolation, and Durability (so-called ACID properties) without any further work (nor even needing to understand what useful guarantees these basic properties ensure).

@@

\subsection{Simplicity} 
When a program doesn't quite do what is wanted, it is tempting to add more features or variables, or to treat the problem as an ``exception'' and program around it --- which inserts more code and, almost certainly, more bugs. This way lies over-fitting, a problem familiar from statistics (and machine learning). Programs can be made over-complex and they can then do anything; an over-complex program may seem correct by accident. Instead, the hallmarks of good science are that of parsimony and simplicity; if a simple program can do what is needed it is more likely to be correct. A simpler program is easier to prove correct, easier to program, and easier to debug. A special case of needing simplicity is when fixing bugs: instead of fixing bugs one at a time, one should be fixing the \emph{reasons\/} why the bugs have happened. Generally, when bugs are fixed, programmers should determine \emph{why\/} the bugs occurred, and thence repair the program more strategically.

\subsection{Compliance with standards}

@@

\subsection{Effective multidisciplinary teamwork}

@@

\subsection{Continuous Professional Development (CPD)}

@@

\subsection{Security and other factors}
Of course, there are many other factors to be considered for the professional development of critical code, such as using appropriate methods to ensure cybersecurity \cite{security-engineering,cyber-cacm}, particularly while also being able to up- and download secure updates.

For pandemic modeling specifically, understanding the limitations of numerical methods (in particular, how numerical methods are affected by the choice of programming language and style of programming) is critical.\footnote{{For example, code from one of the surveyed paper \cite{example-numerical-error} uses literal numbers at far too high a precision for the chosen language to be able to represent correctly (conformant implementations use IEEE 754 double precision 64-bit floating point). Such an error typically has an undefined impact on results, and unfortunately is easy to overlook as the program almost certainly ignores the error when running. The error belies misunderstandings in programming which may have wider effects, such as consequences of relying on the precision being higher than it is.}} Hamming \cite{hamming} is considered a classic, but there is a huge choice available.

@@

\subsection{Software is a human activity}

@@

Following the Dunning-Kruger Effect \cite{dunning-kruger,dunning-kruger-numeracy}, programmers over-estimate their programming skills because they do not have the skills to recognize their lack of knowledge --- in the present case, knowledge of basic Software Engineering.

Dunning and Kruger go on to say,

@@

\section{Code, data, and publication}
\globalLabel{on-code-data-publication}

@@

\subsection{When magic numbers become magic code}
Data often controls the flow of code. For example, data summarizing patients may include their gender, but the program processes males and females differently. Then data becomes code.

@@

\subsection{When data is code}
{Many computer programs blur the simplistic code/data distinctions deliberately, to create virtual machines. Data is then run on the virtual machine as program. Many programs provide standard features to do this, such as LISP's and JavaScript's \texttt{eval} functions. Henderson's book \cite{henderson} builds an elegant Pascal program to run \emph{any\/} LISP program as data, and then shows that the LISP program can run itself running other programs, so it is now its own code and \emph{its\/} data --- despite being purely data to the Pascal program. There are numerous advantages to doing this, including: the Pascal program is not just reading data, but structured data that must conform to the rules of LISP; the LISP running itself runs faster than the original Pascal running LISP, even though the Pascal virtual machine is still doing it in the recursive case; LISP is a much more powerful language than Pascal, so a virtual machine can be used to escape the barriers of a limited implementation such as Pascal. In short, any distinctions between code and data are impossible to maintain.}

@@

\subsection{Exploiting code as data for more reliable science}
In the present paper, we knowingly built on this blur between data and code, a special case of \RAPstarp. However, what we did was not unusual except in our explicit and rigorous approach to managing and summarizing data reliably in the paper.

@@

\subsection{When data is text: Exploiting code for reliable publication}

@@

\subsection{Data and polynomials used in the paper}
The data and polynomials used in the main paper's section \ref{over-fit}, and illustrated in the paper's figure \ref{fig-overfit}, is presented below, as generated in \LaTeX\ by the same \emph{Mathematica\/} notebook that generated section \ref{over-fit}. Of course, in the paper itself, the specific data was not necessary to make the point, but if there is any need to replicate it or otherwise scrutinize the arguments in the paper --- as there would be for more complex arguments in typical papers --- the data used and results are shown below.

The table below is a \LaTeX\ table generated by \emph{Mathematica}, and is exactly the data used in the paper.\footnote{Barring coding or other errors of course, which here we checked against manually by comparing this table typeset in \LaTeX\ against the raw data in the original \emph{Mathematica\/} data table, but in general might better be done by an automatic round trip --- though that would not easily spot \LaTeX\ errors.} In a similar way, data and computed results could be presented and made available in other scientific papers.

@@

\subsection{Comparing conventional and RAP approaches}
\globalLabel{RAP-diagrams}
The similarities and differences between the conventional copy-and-paste approach to filling in data and diagrams in publications, the improved systematic RAP process, and using notebooks (such as \emph{Mathematica\/} or Jupyter), are illustrated in the sequence of schematics of figures \ref{fig:basic}, \ref{fig:excel}, \ref{fig:rap}, and \ref{fig:notebook}.

@@

\section{The Spiegelhalter trustworthiness questions}
\globalLabel{supplementary-Spiegelhalter-section}

@@

\def\question#1{\item \emph{#1\/}}
\def\sequestion#1{\begin{itemize}\raggedright\item[$\blacktriangleright$] \textbf{#1}\end{itemize}}

\subsection{How trustworthy are the numbers?}
\newcounter{resumeCounter}

@@

\sequestion{If the claims are exploratory, weaker standards of coding can be used; if the claims are a basis for critical decisions, then there should be evidence of using appropriate Software Engineering (such as defensive programming) to provide appropriate confidence in the results claimed.}
\setcounter{resumeCounter}{\value{enumi}}
\end{enumerate}

\subsection{How trustworthy is the source?}
\begin{enumerate}
\setcounter{enumi}{\value{resumeCounter}}
\question{How reliable is the source of the story?} Consider the possibility of a biased source with conflicts of interest, and check publication is independently peer-reviewed. Ask yourself, `Why does this source want me to hear this story?'

@@

\setcounter{resumeCounter}{\value{enumi}}
\end{enumerate}

\subsection{How trustworthy is the interpretation?}
\begin{enumerate}
\setcounter{enumi}{\value{resumeCounter}}
\question{How does the claim fit with what else is known?} Consider the context, appropriate comparators, including historical data, and what other studies have shown, ideally in a meta-analysis.

@@

\section{A pilot survey of computational science}
\globalLabel{survey-section}

@@

The \RAPstar\ approach cannot enforce the truth of such as summary, but it (and the opportunity to read and review it easily) very significantly increases the chances that the summary is correct and up to date. For example, if refactoring leads to an option being deleted, then it will also disappear \emph{with no further work\/} from the table above. Also, since the summary was proof-read at the same time as proof-reading this \supplement, following \RAPstar\ also increases the chances that any errors or functionality omissions or issues in the makefile have been detected and corrected.

\subsection{Selected journal case studies}
There are numerous case studies to be made from journals and their relation to code and data. Here, we select just two.

\subsubsection{\emph{The Lancet\/}}

@@

\subsubsection{\emph{Journal of Vascular Surgery}}\label{vascular-surgery-note}
In 2020 the \emph{Journal of Vascular Surgery\/} published a research paper \cite{jvs1}, which had to be retracted on ethical grounds \cite{jvs2,jvs3}: it was a na\"\i ve study and the editorial process was unaware of digital norms. Notably, the paper fails to provide access to its anonymized data (with or without qualification), and fails to define the data anonymization algorithm, and also fails to even mention the code that it developed and used to perform its study. The journal's data policy is itself very weak (the authors ``should consider'' including a footnote to offer limited access to the data) and, despite basic statistics policies, it has no policy at all for code (see  section \ref{supplementary-journal-policies-section}). Ironically, the retracted article \cite{jvs1} is still online (as of August 2020) with no reference to any editorial statement to the effect that it has been retracted, despite this being trivial --- and necessary --- to achieve in the widely-accessed online medium.

@@

\subsection{Pilot paper sample}
A sample of \plural{\dataN}{recent paper} covering a broad range of science were sampled from the leading journals \journalBreakdown.

@@

%\newcount \total
%\total=\dataN
%\multiply \total by 4 % three reviewers, one editor
%\advance \total by \countAuthors
%\the\total\
%leading scientists assuming authors, 1 editor, and 3 reviewers per paper, but not counting acknowledged colleagues et al, covering  (plus appendices and supporting material) in total.
%
\caption{Overview of the peer-reviewed paper sample. (Convenient copy of main paper's table \ref{table-overview}.)}
\label{supplement-table-overview}
\end{table*}

@@

\caption{Summary of survey results. (Convenient copy of main paper's table \ref{table-summary}.)}
\label{supplement-table-summary}
\end{table*}

@@

\subsection{Summary of results}
\label{summary-of-results}
The sample selection criteria necessarily identified scientific research with Software Engineering contributions.

@@

\subsubsection{Current code policies of sampled journals}
\globalLabel{supplementary-journal-policies-section}

It is noteworthy that none of the journals sampled permit any reliable style of managing data in published papers, such as described above in sections \ref{detecting-against-error} and \ref{restrictive-policies}. The main paper, section \ref{mentionPLOSandIEEE}, additionally mentions \emph{PLOS ONE\/} and \emph{IEEE Transactions on Software Engineering}.

For all the surveyed papers that had accessible code, the code included explicit (and relevant) data that was not archived \emph{as\/} data in the journal repositories.

Note that journal policies relevant to the survey were first accessed on 29 July 2020, close after the period covered by the pilot survey, thus presumably fairly closely reflecting the policies in use for the papers in the survey on the dates when they were each submitted. Unfortunately for the survey, the journals do not make clear what exact policies were applied to each paper when the papers were submitted (on the other hand, the survey shows that policies are not rigorously enforced).
\\

\newdimen\tabwidth
\tabwidth=\textwidth
\advance \tabwidth by -1em
\def\specifyurl{\vskip 0mm\noindent\begingroup\small} 
\def\specifyaccess#1{\endgroup\par\noindent
\begin{tabular}{|p{\tabwidth}@{}}#1
\end{tabular}\vskip .3cm
}

\noindent\textbf{Extract from \emph{Royal Society Open Science\/} author guidelines} \\ 
{\sf``It is a condition of publication that authors make the primary data, materials (such as statistical tools, protocols, software) and code publicly available. These must be provided at the point of submission for our Editors and reviewers for peer-review, and then made publicly available at acceptance. [\ldots] As a minimum, sufficient information and data are required to allow others to replicate all study findings reported in the article. Data and code should be deposited in a form that will allow maximum reuse. As part of our open data policy, we ask that data and code are hosted in a public, recognized repository, with an open licence (CC0 or CC-BY) clearly visible on the landing page of your dataset.''}\specifyurl\url{royalsociety.org/journals/authors/author-guidelines/#data}\specifyaccess{Since first accessed 29 July 2020, the policy has been revised (undated, accessed 2 February 2022) but retains the same principles; full policy now available via a DOI \cite{RS-policy}. The policy still retains an emphasis on data accessibility, and continues a lack of awareness that code and data are equivalent and often mixed (see section \ref{on-code-data-publication}).}
